---
title: "APA Style Correlation Tables in R"
output:
  html_document:
    df_print: paged
---

We will create the correlation matrix using `Hmisc::rcorr` to get the correlations along with the p-values and `knitr::kable` to print the table.
Using R-Studio, we can knit directly to word and format it the way we like.
```{r, warning=F, message=F}
library(Hmisc) # For rcorr
library(knitr) # Pretty print table
```

The general format for each of the cell is to display the correlation, along with stars to indicate the significance level.
We want to apply a function (Pasting `*`) for each cell that is below a specific level.

### Helper Function
First we define a function that allows us to apply a function to each cell that satisfies a property.
There may already be a function for this, but I could not find it.
```{r}
#' Applies a function `f` to each cell of a data frame `mat` if the corresponding cell in `p` is TRUE
#'
#' @param mat A matrix or data frame
#' @param p A matrix with the same dimension as `mat`
#' @param f A function to apply
#' @return `mat` with `f` applied to each cell where `p` is TRUE.
#' @examples
#' x <- rbind(c(1,2,3), c(4,5,6), c(7,8,9))
#' apply_if(x, upper.tri(x), function(x) x + 5)
apply_if <- function(mat, p, f) {
  for(i in 1:length(mat[,1])) {
    for(j in 1:length(mat[i,])) {
      if(is.na(p[i,j]) || p[i,j])  { 
        mat[i,j] = f(mat[i,j])
      }
    }
  }
  mat
}
```

### APA Correlation Matrix
Now we can define a function to return the matrix with the stars appended to significant correlations.
The row names are kept with the index appended.
The column names are replaced with the index.
```{r}
#' @param mat an rcorr object or a double matrix
#' @param corrtype is either pearson or spearman. Will be passed into Hmsic::rcorr if mat is not already an rcorr object
#' @return `mat` with stars appended for each level of significants (p < 0.05, p < 0.01, p < 0.001)
apaCorr <- function(mat, corrtype = "pearson") {
  matCorr <- mat
  if (class(matCorr) != "rcorr") {
    matCorr <- rcorr(mat, type = corrtype)
  }

  # Add one star for each p < 0.05, 0.01, 0.001
  stars <- apply_if(round(matCorr$r, 2), matCorr$P < 0.05, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.01, function(x) paste0(x, "*"))
  stars <- apply_if(stars, matCorr$P < 0.001, function(x) paste0(x, "*"))
  # Put - on diagonal and blank on upper diagonal
  stars[upper.tri(stars, diag = T)] <- "-"
  stars[upper.tri(stars, diag = F)] <- ""
  n <- length(stars[1,])
  colnames(stars) <- 1:n
  # Remove _ and convert to title case
  row.names(stars) <- tools::toTitleCase(sapply(row.names(stars), gsub, pattern="_", replacement = " "))
  # Add index number to row names
  row.names(stars) <- paste(paste0(1:n,"."), row.names(stars))
  stars
}
```

### Example
Now we can use `apaCorr` and `kable` to pretty print the correlation table in a format similar to APA.
```{r}
irisStars <- apaCorr(as.matrix(mtcars), corrtype = "pearson")

kable(irisStars, format = "markdown")
```


